### 2020-04-14 thesis/content/chapter-graph-lta.tex
improved intro and conclusion
corrections

In the previous chapter an introduction to two separate fields of research was given:
\begin{enumerate*}
	\item \Acf{lta},
	\item \Acf{gcr}
\end{enumerate*}.
In this chapter we will combine them and define an extension of \ac{lta} to the \ac{gcr} problem.
This will be done in three steps:
\begin{enumerate}
	\item We begin with a formal definition of what actually constitutes an \ac{lta} method as opposed to non-\acs{lta} methods.
	\item Using this definition, we will see that an \ac{svm} that uses one of the previously described graph embedding approaches can be interpreted as an \ac{lta} variant under certain conditions.
	\item In the last step we look at the previously described \ac{gcnn} architectures and formalize their relation to \ac{lta}.
\end{enumerate}

@@

		In the existing unstructured \ac{lta} approaches the decomposition function is implicitly defined as $\psi(G) \coloneqq {\{ G[v_{c_i}] \}}_{v_{c_i} \in \mathcal{V}_G}$ since each vertex $v_{c_i}$ corresponds to an interpretable constituent $c_i$ by definition.
		For structured data however, a split into individual vertices is typically not appropriate.
		Molecular graphs from chemical datasets for example are meaningfully characterized by the presence of so-called \textit{functional groups} consisting of multiple bonded atoms while a characterization on the level of individual atoms is generally less meaningful~\cite{McNaught1997}.
	\item \textbf{Local evaluation:}
		The constituents $c_{G, i} \in \psi(G)$ must be evaluated via some function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}$.
		This \textit{evaluation function} assigns a prediction $\hat{y}_{G, i} \in \mathcal{Y}$ and a weight $w_{G, i} \in \mathbb{R}$ to each constituent.
		A constituent's weight $w_{G, i}$ can intuitively be interpreted as a measure of the confidence that the local prediction $\hat{y}_{G, i}$ is indicative of the composition's global target score $y_G$.
		Learning local predictions and weights for all possible constituents is called the \textit{disaggregation problem}.

@@

		By weighting the constituents, an \ac{lta} model can reduce the relevance or even ignore constituents that turn out to be irrelevant for the compositions target score $y_G$.

@@

		Lastly a \textit{weighted aggregation function} $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ must be applied.

@@

				Unlike $\wmean$ it can also be applied to score domains $\mathcal{Y}$ without a multiplication operator, e.g.\ sets of discrete classes.
			\item Alternatively an unweighted aggregation function like $\min$, $\max$, $\mean$ or \ac{owa}\footnote{

@@

\begin{defn}

@@

Note that every model $h: \mathcal{G} \to \mathcal{Y}$ has a trivial recursive \ac{lta} formulation by choosing $\psi(G) = \{ G \}$, $f(G) = (h(G), 1)$ and an arbitrary weighted aggregation function $\mathcal{A}$.

@@

However since there is no commonly accepted formal criterion to decide whether a model's decisions are explainable~\cite{Lipton2018}, we do not attempt to strictly distinguish between \ac{lta} and non-\acs{lta} methods.
Instead the notion of \ac{lta} formulations should be seen as way to identify how ``\acs{lta}-like'' a model is:

@@

The more ``\acs{lta}-like'' a given model is, the stronger its bias towards locally explainable predictions, which in turn reduces the potential expressive power of the model.
\citet{Gilpin2018} describe this trade-off between explainability and expressive power in more detail.
However, when considering problem domains in which the true composition scores $y_G$ are accurately described by an \acs{lta}-like generative process, a less expressive \ac{lta}-like model could generalize better than a more expressive non-\ac{lta} model.
This idea is captured by the so-called \textit{\ac{lta} assumption}.
\begin{defn}\label{defn:ltag:lta-assumption}
	A problem domain $\mathcal{D}$ satisfies the \textit{\ac{lta} assumption} iff.\ there is an \ac{lta} method which produces models with an equal or lower out-of-sample error than the models produced by non-\acs{lta} methods for most training samples $\mathcal{D}_{\mathrm{train}} \subseteq \mathcal{D}$.
\end{defn}
Due to the fuzziness of the class of \ac{lta} methods, the \ac{lta} assumption is naturally also a fuzzy concept.
Nonetheless evidence for its truthiness in a given domain $\mathcal{D}$ can be empirically obtained by comparing candidate \ac{lta} methods with the best known non-\ac{lta} method for $\mathcal{D}$, assuming that some cut-off condition for the required ``\ac{lta}-ness'' of an \ac{lta} method is agreed upon.

@@

Based on the general definition of \ac{lta} from the last section, we will now see to what extent the \ac{gcr} methods described in \cref{sec:related:gcr} can be interpreted as \ac{lta} instances.
This section explores the relation between \acp{svm} that use graph embeddings and \ac{lta}.
In \cref{sec:related:gcr:embed,sec:related:gcr:kernel} three different ways to map a given graph $G$ to a vector $\varphi(G) \in \mathbb{R}^d$ were described:
\begin{enumerate*}
	\item Fingerprint embeddings,
	\item skip-gram inspired embeddings and
	\item kernel embeddings
\end{enumerate*}.
One common way to solve the \ac{gcr} problem via those embedding vectors is to train an \ac{svm} on them.
We will now see that \acp{svm} can be interpreted as \ac{lta} models if they are trained on so-called \acp*{sce}.
\begin{defn}
	Given a a multiset $A = \ldblbrace \underbrace{a_1, \dots, a_1}_{\gamma_A(a_1)\text{ times}},\allowbreak \dots, \underbrace{a_n, \dots, a_n}_{\gamma_A(a_n)\text{ times}} \rdblbrace \subseteq D$, the so-called \textit{multiplicity function} $\gamma_A: D \to \mathbb{N}_0$ maps each element of the domain $D$ to its multiplicity in $A$ with $\gamma_A(x) = 0 \Leftrightarrow x \notin A$.
\end{defn}
\begin{defn}\label{defn:ltag:substruct-embedding}
	A graph embedding $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is called a \ac{sce} iff.\ %
	there are decomposition functions $\psi_{\varphi, i}: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and multiplicity functions $\gamma_{\varphi, i}: \mathcal{G} \to \mathbb{N}_0$ for all embedding components $i \in [d]$ s.t.\ %
	$\forall G \in \mathcal{G}, i \in [d]: \varphi(G)[i] = \sum_{c \in \psi_{\varphi, i}(G)} \gamma_{\varphi, i}(c)$ where $\gamma_{\varphi, i}$ is the multiplicity function of a multiset of the constituents $\psi_{\varphi, i}(G)$.
	We call $\psi_{\varphi, i}$ an \textit{underlying decomposition} of the $i$-th component of the embedding $\varphi$.
	Similarly the joint decomposition $\psi_{\varphi}(G) \coloneqq \bigcup_{i=1}^d \psi_{\varphi,i}(G)$ is an \textit{underlying decomposition} of $\varphi$.
\end{defn}
Intuitively \cref{defn:ltag:substruct-embedding} states that the value of each \ac{sce} component must be derived from the number of constituents produced by some decomposition function where each constituent might be counted multiple times.
Based on this requirement we now proof the main theorem which shows the relation between \acp{svm} and \ac{lta}.

\begin{thm}\label{thm:ltag:svm-ltag-formulation}
	A binary \ac{svm} graph classifier $h$ that applies a \ac{sce} $\varphi$ to its inputs has an \ac{lta} formulation that uses an underlying decomposition $\psi_{\varphi}$ of $\varphi$.
\end{thm}
\begin{proof}
	Let $h: \mathcal{G} \to {\{ -1, +1 \}}$ be a binary graph classifier expressed as $h = h_{\mathrm{\acs*{svm}}} \circ \varphi$ where $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is an \ac{sce} and $h_{\mathrm{\acs*{svm}}}: \mathbb{R}^{d} \to {\{ -1, +1 \}}$ a standard \ac{svm} classifier.
	Additionally, let $\psi_{\varphi}$ be some underlying decomposition of $\varphi$ with ${\{ \gamma_{\varphi, i} \}}_{i=1}^d$ being the corresponding multiplicity functions.

@@

	To complete the proof it now remains to show that $f_h$ and $\mathcal{A}_h$ are in fact a local evaluation function and a weighted aggregation function respectively.
	To see that $f_h$ perform local evaluation, note that $f_h(c_t) \coloneqq (\sgn z_t, {|z_t|})$ with $z_t \coloneqq \sum_{i=1}^{d} \gamma_{\varphi,i}(c_t) \beta_i$ only depends on the shared multiplicity functions $\gamma_{\varphi,i}$, the constants $\beta_i$ and the constituent $c_t$;
	apart from $c_t$ no other information from the input graph $G$ is required by $f_h$ which makes it a local evaluation function.
	To see why $\mathcal{A}_h$ satisfies \cref{defn:ltag:weighted-agg}, note that it inherits the idempotency property from $\wmaj$ because $\mathcal{A}_h$ ignores the bias ``pseudo-constituent'' $(y_b, w_b)$ for all threshold weights $\max w_t \geq \eta > w_b$, similarly the zero invariance of $\wmaj$ is also directly inherited.
	This concludes the proof.
\end{proof}

@@

To see what \cref{thm:ltag:svm-ltag-formulation} implies in practice, we will now apply it to the graph embedding approaches described in \cref{sec:related:gcr:embed,sec:related:gcr:kernel}:
\begin{enumerate}[label=\textbf{\arabic*.},ref={\arabic*}]
	\item \textbf{Fingerprint embeddings:}\label{itm:ltag:fingerprint-lta-formulation}
		Here the $i$-th embedding component represent the number of occurrences of a substructure $S_i$, i.e.\ ${\varphi}_{\mathrm{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G)$ (see \cfullref{defn:related:wl-count-detect}).
		Such a fingerprint embedding naturally is an \ac{sce} with the decomposition functions $\psi_{\mathrm{FP}, i}(G) \coloneqq \{ G[\set(s)]\, |\, s \in \mathcal{V}_G^{*} \land G[s] \equiv S_i \}$ and the multiplicity functions $\gamma_{\mathrm{FP}, i}(c) \coloneqq \mathbbm{1}[c \simeq S_i]$ since those functions produce the subgraphs that are counted by ${\varphi}_{\mathrm{FP}}(G)[i]$ with multiplicity $1$. % chktex 21

		Under the assumption that the substructure patterns $S_i$ are chosen s.t.\ the their instances $c \in \psi_{\mathrm{FP}, i}(G)$ are nontrivial interpretable constituents, the underlying joint decomposition $\psi_{\mathrm{FP}}(G) \coloneqq \bigcup_{i=1}^d \psi_{\mathrm{FP}, i}(G)$ must also be nontrivial and interpretable.
		Thus, by \cref{thm:ltag:svm-ltag-formulation}, \acp{svm} trained on fingerprint embeddings are \ac{lta} models that produce locally explainable predictions.
		\badgebox[t_green]{\acs*{lta}}
	\item \textbf{\texttt{graph2vec}:}
		The skip-gram inspired \texttt{graph2vec} embedding produces vectors $\varphi(G) \in \mathbb{R}^d$ whose individual components do not have a clear interpretation.
		\texttt{graph2vec} embeds graphs that have common 1-\acs{wl} colors closer to each other than graphs that do not share the same colors (see \cfullref{eq:related:graph2vec}).
		The resulting component values can be arbitrary reals, therefore this approach is not an \ac{sce} which in turn implies that it does not have an \ac{lta} formulation as in \cref{thm:ltag:svm-ltag-formulation}.
		\badgebox[t_red]{non-\acs*{lta}}
	\item \textbf{\ac{wl} subtree kernel:}\label{itm:ltag:wlst-lta-formulation}
		The individual components of the \ac{wl} subtree kernel's embedding $\varphi_{\mathrm{ST}}$ represent the number of occurrences of a color $\kappa_i \in \mathcal{C}$, i.e.\ $\varphi_{\mathrm{ST}}(G)[i, t] \coloneqq \mathit{dist}_{\chi_{G,1}^{(t)}}(\kappa_i)$ as defined in \cfullref[ on ]{eq:related:wl-subtree-kernel}\footnote{
			To avoid confusion $\kappa_i \in \mathcal{C}$ is used for colors and $c_j \in \mathcal{G}$ for constituents in this context.
		}.
		The occurrence of a color $\kappa_i$ at some vertex $v_j \in \mathcal{V}_G$ in the $t$-th refinement step (i.e.\ $\chi_{G,1}^{(t)}(v_j) = \kappa_i$) in turn implies that the \ac{bfs} tree $S_{j,t} \coloneqq \text{\acs*{bfs}}(G, v_j, t)$ of depth $t$ rooted at $v_j$ is isomorphic to some \ac{wl} color subtree $S_{\kappa_i}$ (i.e.\ $S_{j,t} \simeq S_{\kappa_i}$).
		This relation between \ac{wl} colors $\kappa_i$ and subtrees $S_{\kappa_i}$ was illustrated in \cfullref[ on ]{fig:related:wl-subtree}.

		To show that $\varphi_{\mathrm{ST}}$ is an \ac{sce}, we have to define decomposition functions $\psi_{\varphi, i, t}$ and multiplicity functions $\gamma_{\varphi, i, t}$ s.t.\ $\varphi(G)[i, t] = \sum_{c \in \psi_{\varphi, i, t}(G)} \gamma_{\varphi, i, t}(c)$.
		Since $\varphi(G)[i, t]$ counts the occurrences $S_{j,t}$ of a subtree $S_{\kappa_i}$ in $G$, the \ac{sce} requirement would be trivially satisfied if each $S_{j,t}$ were a constituent with multiplicity $1$.
		This intuition is however not quite correct because the constituents of a graph $G$ must be induced subgraphs of $G$, not \ac{bfs} subtrees $S_{j, t}$ of $G$.
		\begin{figure}[t]
			\centering
			\includegraphics[width=\linewidth]{gfx/graph-lta/wl1-constituents.pdf}
			\caption[The constituents implied by the \ac{wl} subtree kernel embedding for an example graph.]{
				The constituents $c_{j,t}$ implied by the \ac{wl} subtree embedding vector $\varphi(G) = (\underbrace{\textcolor{t_blue}{4}, \textcolor{t_red}{0}, \textcolor{t_green}{0}, \textcolor{t_darkyellow}{0}}_{t = 0}, \underbrace{\textcolor{t_blue}{0}, \textcolor{t_red}{2}, \textcolor{t_green}{1}, \textcolor{t_darkyellow}{1}}_{t = 1})$. % chktex 25
				The subtrees $S_{j,t}$ that span each constituent $c_{j,t}$ are visualized by the colored arrows that originate from the colored root vertices.
				Because $S_{1,1}$ and $S_{2,1}$ span the same constituent $c = c_{1,1} = c_{2,1} = G[\{ v_1, v_2, v_3 \}]$, its multiplicity is $\gamma_{\varphi, \textcolor{t_red}{\texttt{B}}, 1}(c) = 2$; % chktex 25
				the multiplicity of all other constituents is $1$.
			}\label{fig:ltag:wl1-constituents}
		\end{figure}

		To fix the previous intuition we ``convert'' the subtrees $S_{j, t}$ into proper subgraph constituents via $c_{j,t} \coloneqq G[\mathcal{V}_{S_{j,t}}]$, i.e.\ the subgraphs that are spanned by the subtrees.
		The resulting decomposition functions are $\psi_{\varphi, i, t}(G) \coloneqq \{ c_{j,t}\, |\, {v_j \in \mathcal{V}_G}\,\land {S_{j,t} \simeq S_{\kappa_i}} \}$. % chktex 21
		As illustrated in \cref{fig:ltag:wl1-constituents}, the number of distinct constituents $c_{j,t}$ might be smaller than the number of \ac{bfs} subtrees $S_{j,t}$ because two distinct subtrees might span the same set of vertices.
		To fix this discrepancy between the subtree occurrence count (which equals $\varphi(G)[i, t]$) and the number of constituents, the multiplicities $\gamma_{\varphi, i, t}(c)$ of the constituents $c$ have to correspond to the number of \ac{bfs} subtrees they were spanned up by, i.e.\
		\begin{align}
			&\gamma_{\varphi, i, t}(c) \coloneqq {\left|\{ S = \text{\acs*{bfs}}(c, v_{\mathrm{root}}, t)\, |\, v_{\mathrm{root}} \in \mathcal{V}_c \land S \simeq S_{\kappa_i} \land \mathit{complete}(S) \}\right|}\label{eq:ltag:wlst-constituent-subtree-count} \\
			&\text{with }\mathit{complete}(S) \Leftrightarrow \forall\ \text{leaf nodes $v_{\mathrm{leaf}}$ of the tree $S$}: {|\Gamma_c(v_{\mathrm{leaf}})|} = {|\Gamma_G(v_{\mathrm{leaf}})|}
			\text{.} \nonumber
		\end{align}
		The purpose of the $\mathit{complete}(S)$ condition is to only count the \ac{bfs} subtrees of the constituent $c$ that are also \ac{bfs} subtrees of the complete graph $G$.
		To see why this is required, note that in \cref{fig:ltag:wl1-constituents} the $c_{1,1}$/$c_{2,1}$ constituent contains the color subtree $S_{\textcolor{t_red}{\texttt{B}}}$ three times (rooted at $v_1$, $v_2$ and $v_3$) even though $G$ only contains it twice (rooted at $v_1$ and $v_2$). % chktex 25
		Also note that, since $\gamma_{\varphi, i, t}(c)$ must only depend on a given constituent $c$ and not on the graph $G$ it was decomposed from, the degree information $|\Gamma_G(v_{\mathrm{leaf}})|$ of the constituent vertices $v \in \mathcal{V}_c$ is assumed to be statically encoded in the labels $l_c[v]$ or feature vectors $x_c[v]$.

>>>>

In the previous chapter an introduction to two separate fields of research was given:
\begin{enumerate*}[label={\circled{\small\arabic*}}]
	\item \Acf{lta}.
	\item \Acf{gcr}.
\end{enumerate*}
In this chapter we combine them and define an extension of \ac{lta} to the \ac{gcr} problem.
This will be done in three steps:
\begin{enumerate}
	\item We begin with a formal definition of what actually constitutes an \ac{lta} method as opposed to non-\acs{lta} methods.
	\item Using this definition, we will see that an \ac{svm} that uses one of the previously described graph embedding approaches can be interpreted as an \ac{lta} variant under certain conditions.
	\item In the last step we look at the previously described \ac{gcnn} architectures and formalize their relation to \ac{lta}.
\end{enumerate}

@@

		In the existing unstructured \ac{lta} approaches the decomposition function is implicitly defined as $\psi(G) \coloneqq {\{ G[v_{c_i}] \}}_{v_{c_i} \in \mathcal{V}_G}$ since each vertex $v_{c_i}$ corresponds to an interpretable constituent $c_i$ by definition.
		For structured data however, a split into individual vertices is typically not appropriate.
		Molecular graphs from chemical datasets for example are meaningfully characterized by the presence of so-called \textit{functional groups} consisting of multiple bonded atoms while a characterization on the level of individual atoms is generally less meaningful~\cite{McNaught1997}.
	\item \textbf{Local evaluation:}
		The constituents $c_{G, i} \in \psi(G)$ must be evaluated via some function $f: \mathcal{G} \to \mathcal{Y} \times \mathbb{R}_{\geq 0}$.
		This \textit{evaluation function} assigns a prediction $\hat{y}_{G, i} \in \mathcal{Y}$ and a weight $w_{G, i} \in \mathbb{R}_{\geq 0}$ to each constituent.
		A constituent's weight $w_{G, i}$ can intuitively be interpreted as a measure of the confidence that the local prediction $\hat{y}_{G, i}$ is indicative of the composition's global target score $y_G$.
		Learning local predictions and weights for all possible constituents is called the \textit{disaggregation problem}.

@@

		By weighting the constituents, an \ac{lta} model can reduce the relevance or even ignore constituents that turn out to be irrelevant for the composition's target score $y_G$.

@@

		Lastly, a \textit{weighted aggregation function} $\mathcal{A}: {(\mathcal{Y} \times \mathbb{R}_{\geq 0})}^{*} \to \mathcal{Y}$ must be applied.

@@

				Unlike $\wmean$, it can also be applied to score domains $\mathcal{Y}$ without a multiplication operator, e.g.\ sets of discrete classes.
			\item Alternatively, an unweighted aggregation function like $\min$, $\max$, $\mean$ or \ac{owa}\footnote{

@@

\begin{defn}\label{defn:ltag:formulation}

@@

Note that every model $h: \mathcal{G} \to \mathcal{Y}$ has a trivial recursive \ac{lta} formulation by choosing $\psi(G) = \{ G \}$, $f(G) = (h(G), \eta)$ and an arbitrary weighted aggregation function $\mathcal{A}$.

@@

However, since there is no commonly accepted formal criterion to decide whether a model's decisions are explainable~\cite{Lipton2018}, we do not attempt to strictly distinguish between \ac{lta} and non-\acs{lta} methods.
Instead, the notion of \ac{lta} formulations should be seen as way to identify how ``\acs{lta}-like'' a model is.

@@

The more ``\acs{lta}-like'' a given model is, the stronger its bias towards locally explainable predictions, which in turn reduces the potential expressive power of the model.
\citet{Gilpin2018} describe this trade-off between explainability and expressive power in more detail.
However, when considering problem domains in which the true composition scores $y_G$ are accurately described by an \acs{lta}-like generative process, a less expressive \ac{lta}-like model could generalize better than a more expressive non-\ac{lta} model.
This idea is captured by the so-called \textit{\ac{lta} assumption}.
\begin{defn}\label{defn:ltag:lta-assumption}
	A problem domain $\mathcal{D}$ satisfies the \textit{\ac{lta} assumption} iff.\ there is an \ac{lta} learner which produces models with an equal or lower out-of-sample error than the models produced by non-\acs{lta} learners for most training samples $\mathcal{D}_{\mathrm{train}} \subseteq \mathcal{D}$.
\end{defn}
Due to the fuzziness of the class of \ac{lta} methods, the \ac{lta} assumption is naturally also a fuzzy concept.
Nonetheless, evidence for its truthiness in a given domain $\mathcal{D}$ can be empirically obtained by comparing candidate \ac{lta} methods with the best known non-\ac{lta} method for $\mathcal{D}$, assuming that some cut-off condition for the required ``\ac{lta}-ness'', i.e.\ constituent interpretability, of an \ac{lta} method is agreed upon.

@@

Based on the general definition of \ac{lta} from the last section, we will now see to what extent the \ac{gcr} methods described in \cref{sec:related:gcr} can be interpreted as \ac{lta} instances.
This section explores the relation between \ac{lta} and \acp{svm} that use graph embeddings.
In \cref{sec:related:gcr:embed,sec:related:gcr:kernel}, three different ways to map a given graph $G$ to a vector $\varphi(G) \in \mathbb{R}^d$ were described:
\begin{enumerate*}[label={\circled{\small\arabic*}}]
	\item Fingerprint embeddings,
	\item skip-gram inspired embeddings and
	\item kernel embeddings
\end{enumerate*}.
One common way to solve the \ac{gcr} problem via those embedding vectors is to train an \ac{svm} on them.
We will now see that \acp{svm} can be interpreted as \ac{lta} models if they are trained on so-called \acp*{sce}.
\begin{defn}
	Given a multiset $A = \ldblbrace \underbrace{a_1, \dots, a_1}_{\gamma_A(a_1)\text{ times}},\allowbreak \dots, \underbrace{a_n, \dots, a_n}_{\gamma_A(a_n)\text{ times}} \rdblbrace \subseteq D$, the so-called \textit{multiplicity function} $\gamma_A: D \to \mathbb{N}_0$ maps each element of the domain $D$ to its multiplicity in $A$, with $\gamma_A(x) = 0 \Leftrightarrow x \notin A$.
\end{defn}
\begin{defn}\label{defn:ltag:substruct-embedding}
	A graph embedding $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is called a \ac{sce} iff.\ %
	there are decomposition functions $\psi_{\varphi, i}: \mathcal{G} \to \mathcal{P}(\mathcal{G})$ and multiplicity functions $\gamma_{\varphi, i}: \mathcal{G} \to \mathbb{N}_0$ for all embedding components $i \in [d]$ s.t.\ %
	$\forall G \in \mathcal{G}, i \in [d]: \varphi(G)[i] = \sum_{c \in \psi_{\varphi, i}(G)} \gamma_{\varphi, i}(c)$, where $\gamma_{\varphi, i}$ is the multiplicity function of a multiset of the constituents $\psi_{\varphi, i}(G)$.
	We call $\psi_{\varphi, i}$ an \textit{underlying decomposition} of the $i$-th component of the embedding $\varphi$.
	Similarly, the joint decomposition $\psi_{\varphi}(G) \coloneqq \bigcup_{i=1}^d \psi_{\varphi,i}(G)$ is an \textit{underlying decomposition} of $\varphi$.
\end{defn}
Intuitively, \cref{defn:ltag:substruct-embedding} states that the value of each \ac{sce} component must be derived from the number of constituents produced by some decomposition function, where each constituent might be counted multiple times.
Based on this requirement we now proof the main theorem which shows the relation between \acp{svm} and \ac{lta}.

\begin{thm}\label{thm:ltag:svm-ltag-formulation}
	A binary \ac{svm} graph classifier $h$ that applies an \ac{sce} $\varphi$ to its inputs has an \ac{lta} formulation that uses an underlying decomposition $\psi_{\varphi}$ of $\varphi$.
\end{thm}
\begin{proof}
	Let $h: \mathcal{G} \to {\{ -1, +1 \}}$ be a binary graph classifier expressed as $h = h_{\mathrm{\acs*{svm}}} \circ \varphi$, where $\varphi: \mathcal{G} \to \mathbb{N}_0^{d}$ is an \ac{sce} and $h_{\mathrm{\acs*{svm}}}: \mathbb{R}^{d} \to {\{ -1, +1 \}}$ a standard \ac{svm} classifier.
	Additionally, let $\psi_{\varphi}$ be some underlying decomposition of $\varphi$, with ${\{ \gamma_{\varphi, i} \}}_{i=1}^d$ being the corresponding multiplicity functions.

@@

	To complete the proof it now remains to show that $f_h$ and $\mathcal{A}_h$ are in fact a local evaluation function and a weighted aggregation function respectively.
	To see that $f_h$ performs local evaluation, note that $f_h(c_t) \coloneqq (\sgn z_t, {|z_t|})$, with $z_t \coloneqq \sum_{i=1}^{d} \gamma_{\varphi,i}(c_t) \beta_i$ only depending on the shared multiplicity functions $\gamma_{\varphi,i}$, the constants $\beta_i$ and the constituent $c_t$;
	apart from $c_t$ no other information from the input graph $G$ is required by $f_h$ which makes it a local evaluation function.
	To see why $\mathcal{A}_h$ satisfies \cref{defn:ltag:weighted-agg}, note that it inherits the idempotency property from $\wmaj$ because $\mathcal{A}_h$ ignores the bias ``pseudo-constituent'' $(y_b, w_b)$ for all threshold weights $\max w_t \geq \eta > w_b$, similarly the zero invariance of $\wmaj$ is also directly inherited.
	This concludes the proof.
\end{proof}

@@

To see what \cref{thm:ltag:svm-ltag-formulation} implies in practice, we will now apply it to the graph embedding approaches described in \cref{sec:related:gcr:embed,sec:related:gcr:kernel}:
\begin{enumerate}[label=\textbf{\arabic*.},ref={\arabic*}]
	\item \textbf{Fingerprint embeddings:}\label{itm:ltag:fingerprint-lta-formulation}
		Here the $i$-th embedding component represents the number of occurrences of a substructure $S_i$, i.e.\ ${\varphi}_{\mathrm{FP}}(G)[i] \coloneqq \mathit{count}_{S_i}(G)$ (see \cfullref{defn:related:wl-count-detect}).
		Such a fingerprint embedding naturally is an \ac{sce} with the decomposition functions $\psi_{\mathrm{FP}, i}(G) \coloneqq \{ G[\set(s)]\, |\, s \in \mathcal{V}_G^{*} \land G[s] \equiv S_i \}$ and the multiplicity functions $\gamma_{\mathrm{FP}, i}(c) \coloneqq \mathbbm{1}[c \simeq S_i]$ since those decompositions produce the subgraphs that are counted by ${\varphi}_{\mathrm{FP}}(G)[i]$ with multiplicity $1$. % chktex 21

		Under the assumption that the substructure patterns $S_i$ are chosen s.t.\ the their instances $c \in \psi_{\mathrm{FP}, i}(G)$ are nontrivial interpretable constituents, the underlying joint decomposition $\psi_{\mathrm{FP}}(G) \coloneqq \bigcup_{i=1}^d \psi_{\mathrm{FP}, i}(G)$ must also be nontrivial and interpretable.
		Thus, by \cref{thm:ltag:svm-ltag-formulation}, \acp{svm} trained on fingerprint embeddings are \ac{lta} models that produce locally explainable predictions.
		\badgebox[t_green]{\acs*{lta}}
	\item \textbf{\texttt{graph2vec}:}
		The skip-gram inspired \texttt{graph2vec} embedding produces vectors $\varphi(G) \in \mathbb{R}^d$ whose individual components do not have a clear interpretation.
		\texttt{graph2vec} embeds graphs that have common 1-\acs{wl} colors closer to each other than graphs that do not share the same colors (see \cfullref{eq:related:graph2vec}).
		The resulting component values can be arbitrary reals, therefore this approach is not an \ac{sce} which in turn implies that it does not have an \ac{lta} formulation as in \cref{thm:ltag:svm-ltag-formulation}.
		\badgebox[t_red]{non-\acs*{lta}}
	\item \textbf{\ac{wl} subtree kernel:}\label{itm:ltag:wlst-lta-formulation}
		The individual components of the \ac{wl} subtree kernel's embedding $\varphi_{\mathrm{ST}}$ represent the number of occurrences of a color $\kappa_i \in \mathcal{C}$, i.e.\ $\varphi_{\mathrm{ST}}(G)[i, t] \coloneqq \mathit{dist}_{\chi_{G,1}^{(t)}}(\kappa_i)$ as defined in \cfullref[ on ]{eq:related:wl-subtree-kernel}\footnote{
			To avoid confusion, $\kappa_i \in \mathcal{C}$ is used for colors and $c_j \in \mathcal{G}$ for constituents in this context.
		}.
		The occurrence of a color $\kappa_i$ at some vertex $v_j \in \mathcal{V}_G$ in the $t$-th refinement step (i.e.\ $\chi_{G,1}^{(t)}(v_j) = \kappa_i$) in turn implies that the \ac{bfs} tree $S_{j,t} \coloneqq \mathrm{\acs*{bfs}}(G, v_j, t)$ of depth $t$ rooted at $v_j$ is isomorphic to some \ac{wl} color subtree $S_{\kappa_i}$ (i.e.\ $S_{j,t} \simeq S_{\kappa_i}$).
		This relation between \ac{wl} colors $\kappa_i$ and subtrees $S_{\kappa_i}$ was illustrated in \cfullref[ on ]{fig:related:wl-subtree}.

		To show that $\varphi_{\mathrm{ST}}$ is an \ac{sce}, we have to define decomposition functions $\psi_{\varphi, i, t}$ and multiplicity functions $\gamma_{\varphi, i, t}$ s.t.\ $\varphi(G)[i, t] = \sum_{c \in \psi_{\varphi, i, t}(G)} \gamma_{\varphi, i, t}(c)$.
		Since $\varphi(G)[i, t]$ counts the occurrences $S_{j,t}$ of a subtree $S_{\kappa_i}$ in $G$, the \ac{sce} requirement would be trivially satisfied if each $S_{j,t}$ were a constituent with multiplicity $1$.
		This intuition is however not quite correct because the constituents of a graph $G$ must be induced subgraphs of $G$, not \ac{bfs} subtrees $S_{j, t}$ of $G$.
		\begin{figure}[t]
			\centering
			\includegraphics[width=\linewidth]{gfx/graph-lta/wl1-constituents.pdf}
			\caption[The constituents implied by the \ac{wl} subtree kernel embedding for an example graph.]{
				The constituents $c_{j,t}$ implied by the \ac{wl} subtree embedding vector $\varphi(G) = (\underbrace{\textcolor{t_blue}{4}, \textcolor{t_red}{0}, \textcolor{t_green}{0}, \textcolor{t_darkyellow}{0}}_{t = 0}, \underbrace{\textcolor{t_blue}{0}, \textcolor{t_red}{2}, \textcolor{t_green}{1}, \textcolor{t_darkyellow}{1}}_{t = 1})$. % chktex 25
				The subtrees $S_{j,t}$ that span each constituent $c_{j,t}$ are visualized by the colored arrows that originate from the colored root vertices.
				Because $S_{1,1}$ and $S_{2,1}$ span the same constituent $c = c_{1,1} = c_{2,1} = G[\{ v_1, v_2, v_3 \}]$, its multiplicity is $\gamma_{\varphi, \colorlabel{t_red}{\scriptsize B}, 1}(c) = 2$; % chktex 25
				the multiplicity of all other constituents is $1$.
			}\label{fig:ltag:wl1-constituents}
		\end{figure}

		To fix the previous intuition, we ``convert'' the subtrees $S_{j, t}$ into proper subgraph constituents via $c_{j,t} \coloneqq G[\mathcal{V}_{S_{j,t}}]$, i.e.\ the subgraphs that are spanned by the subtrees.
		The resulting decomposition functions are $\psi_{\varphi, i, t}(G) \coloneqq \{ c_{j,t}\, |\, {v_j \in \mathcal{V}_G}\,\land {S_{j,t} \simeq S_{\kappa_i}} \}$. % chktex 21
		As illustrated in \cref{fig:ltag:wl1-constituents}, the number of distinct constituents $c_{j,t}$ might be smaller than the number of \ac{bfs} subtrees $S_{j,t}$ because two distinct subtrees might span the same set of vertices.
		To fix this discrepancy between the subtree occurrence count (which equals $\varphi(G)[i, t]$) and the number of constituents, the multiplicities $\gamma_{\varphi, i, t}(c)$ of the constituents $c$ have to correspond to the number of \ac{bfs} subtrees they were spanned up by, i.e.\
		\begin{align}
			&\gamma_{\varphi, i, t}(c) \coloneqq {\left|\{ S = \mathrm{\acs*{bfs}}(c, v_{\mathrm{root}}, t)\, |\, v_{\mathrm{root}} \in \mathcal{V}_c \land S \simeq S_{\kappa_i} \land \mathit{complete}(S) \}\right|}\label{eq:ltag:wlst-constituent-subtree-count} \\
			&\text{with }\mathit{complete}(S) \Leftrightarrow \forall\ \text{leaf nodes $v_{\mathrm{leaf}}$ of the tree $S$}: {|\Gamma_c(v_{\mathrm{leaf}})|} = {|\Gamma_G(v_{\mathrm{leaf}})|}
			\text{.} \nonumber
		\end{align}
		The purpose of the $\mathit{complete}(S)$ condition is to only count the \ac{bfs} subtrees of the constituent $c$ that are also \ac{bfs} subtrees of the complete graph $G$.
		To see why this is required, note that in \cref{fig:ltag:wl1-constituents} the $c_{1,1}$/$c_{2,1}$ constituent contains the color subtree $S_{\colorlabel{t_red}{\scriptsize B}}$ three times (rooted at $v_1$, $v_2$ and $v_3$) even though $G$ only contains it twice (rooted at $v_1$ and $v_2$). % chktex 25
		Also note that, since $\gamma_{\varphi, i, t}(c)$ must only depend on a given constituent $c$ and not on the graph $G$ it was decomposed from, the degree information $|\Gamma_G(v_{\mathrm{leaf}})|$ of the constituent vertices $v \in \mathcal{V}_c$ is assumed to be statically encoded in the labels $l_c[v]$ or feature vectors $x_c[v]$.
